---
title: "Principal Component Analysis Assignment"
author: "Xi Wang,Natnicha Palakawong Na Ayuthaya,Jaime Vargas Diaz"
date: "11/16/2019"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_float:
      collapsed: no
  pdf_document:
    toc: yes
---

# Function

1. <b>The function must have 2 inputs:</b>
    + <b>a data frame that contains the data set</b>
    + <b>a scaler that denotes the desired percentage of variance preserved in the transformed data set</b>
2. <b>The function must have 3 outputs returned in a list, called number, PCA, and transformed:</b>
    + <b>the minimum number of principal components that is necessary to retain the desired variance
specified by the user</b>
    + <b>the principal components corresponding to previous part, in a matrix</b>
    + <b>the transformed data set in a matrix</b>
    
This function takes two parameters, a dataframe data and a percentage varPercent.
It calculates the minimum number of PCAs necesary to reach the variance percentage, the PCA components, and the new transformed matrix and returns all three in a list.

```{r }
library(psych)
MyPCA<-function(df,scaler){
  newmatrix=data.matrix(df, rownames.force = NA) #transform dataframe to a matrix
  datamean=colMeans(newmatrix) #center the matrix (notice the vectors are row-distributed in a dataframe)
  datacentered=sweep(newmatrix,2,datamean) 
  S=1/(nrow(newmatrix)-1)*(t(datacentered)%*%datacentered) #find the convariance matrix
  Eigenvalue=eigen(S)$value #find eigenvector and eigenvalue
  Eigenvector=eigen(S)$vectors
  var=Eigenvalue/tr(S) #find the fraction of variance each principle components contributes
  a=0
  i=1
  k=0
  while (i<=length(var)) {
    a=a+var[i]
    k=i
    i=i+1
    if (a>= scaler){
      i=length(var)+1}
    }
  number=k #the minimum numbers of principle components required to meet desired variance
  P=Eigenvector[,1:number]  #the principal components corresponding to k
  Y=t(P)%*%t(datacentered) 
  PCA=P
  transformed=t(Y) #transformed dataset in a matrix
  
  
  finalresult<-list(number,PCA,transformed)
  finalresult
  }

```

# Output

<b>
Run your function on the first 5 features of R longley data set with the desired retained variance of 0.98.
Longley data set is a data frame with 7 features related to economics. You will use the following features for
this assignment: GNP.deflator, GNP, Unemployed, Armed.Forces, and Population.
</b>

We will test the function with the first 5 features of R's longley dataset and a desired variance of 0.98.

```{r }
longleysub<-longley[,1:5]
MyPCA(longleysub,0.98)

```

The list returned above shows us the minimum number needed to retain the submitted variance, the principal components, and the new transformed dataset.
